---
title: "Exploratory Data Analysis of Text Corpus"
author: "Ivan Lysiuchenko"
date: "September 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this milestone report we present an overview and exploratory analysis of the three files
we use to build our predictive model. These files represent text corpora extracted from
blogs, news and tweets in English.

## Files overview

To see what the lines of every file look like we use a simple function that can be found
in our Github repository. The minimum, maximum, average line length as well as line counts
are as follows:

```{r message=FALSE, warning=FALSE}
source("../TextFileTools/computeLineStats.R")
fileNames <- c(
    "../materials/datasets/final/en_US/en_US.blogs.txt",
    "../materials/datasets/final/en_US/en_US.news.txt",
    "../materials/datasets/final/en_US/en_US.twitter.txt"
)
fileDescr <- c("Blogs", "News", "Twitter")
linestats <- computeLineStats(fileNames, fileDescr)
linestats
```

We can see that the line length average is similar in the blogs and news texts included in
the corpora. On the opposite, no line belonging to the tweets exceeds 140 characters.

## Loading procedure


